# VegVault Database Structure

Currently, [VegVault]{.vegvault} consists of 31 interconnected tables with 87 fields (variables), which are described in detail below. However, the internal [database structure]{.database} may not be directly relevant to most users, as the [{vaultkeepr}](https://bit.ly/vaultkeepr) R package processes all data to output only the most relevant information in a "[ready-to-analyze]{.data-processing}" format.

::: {.callout-note}
**For Most Users**: If you're primarily interested in using [VegVault]{.vegvault} for research, you may want to start with our [Usage Examples](/website/database_examples.qmd) rather than diving into the technical [database structure]{.database}.
:::

This section provides comprehensive documentation of all tables and their relationships for users who need detailed technical information about the [database architecture]{.database}.

```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("website/database_structure/_01_Description_of_VegVault.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Visualisation/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)

```

<br>

## Metadata Tables

Several tables contain [metadata]{.database} and administrative information that are not directly linked to the scientific data:

- `Authors`: Information about [VegVault]{.vegvault} authors and maintainers, including contact details
- `version_control`: [Database version information]{.database} with descriptions of changes over time
- `sqlite_stat1` & `sqlite_stat4`: [SQLite system tables]{.database} containing [database index statistics]{.database} for query optimization

```{r}
#| label: Authors variables scheme structure
#| tbl-cap: "Column names and types for table `Authors`."
data_comments_authors <-
  tibble::tribble(
    ~column_name, ~description,
    "author_id", "ID of an Author (unique)",
    "author_fullname", "Full name of an Author",
    "author_email", "Contact email of an Author",
    "author_orcid", "ORCID ID of an Author"
  )

get_db_table_data_types(
  con,
  "Authors"
) %>%
  dplyr::full_join(
    data_comments_authors,
    by = "column_name"
  ) %>%
  plot_table()
```

```{r}
#| label: Version Control variables scheme structure
#| tbl-cap: "Column names and types for table `version_control`."
data_comments_version_control <-
  tibble::tribble(
    ~column_name, ~description,
    "id", "ID of a database version (unique)",
    "version", "Version number",
    "update_date", "Date of the creation of that version",
    "changelog", "Text description of main changes in the database"
  )

get_db_table_data_types(
  con,
  "version_control"
) %>%
  dplyr::full_join(
    data_comments_version_control,
    by = "column_name"
  ) %>%
  plot_table()
```

<br>
