---
format:
  html:
    fig-width: 10
  gfm:
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("Technical_overview/_04_section_II_taxa.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Visualisation/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

## Taxa

The **VegVault** database records taxa names derived directly from the primary data sources. Each individual `Taxon` is linked to corresponding `Samples` through the `SampleTaxa` table, ensuring accurate and systematic association between species and their ecological data.  

```{r}
#| label: Taxa scheme structure
knitr::include_graphics(
  here::here(
    "Technical_overview/DB_scheme_visualisation/Taxa.png"
  )
)
``` 

```{r}
#| label: SampleTaxa scheme structure
knitr::include_graphics(
  here::here(
    "Technical_overview/DB_scheme_visualisation/SampleTaxa.png"
  )
)
```

```{r}
#| label: Number of taxa per data source type
n_taxa_per_source_type <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::filter(
    dataset_type == "vegetation_plot"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSourceTypeID"),
    by = "data_source_type_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples"),
    by = "sample_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "SampleTaxa"),
    by = "sample_id"
  ) %>%
  dplyr::distinct(dataset_source_type, taxon_id) %>%
  dplyr::group_by(dataset_source_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

fig_n_taxa_per_source_type <-
  n_taxa_per_source_type %>%
  dplyr::arrange(-N) %>%
  dplyr::mutate(
    dataset_source_type = factor(
      dataset_source_type,
      levels = unique(dataset_source_type)
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = dataset_source_type
    )
  ) +
  ggplot2::scale_y_continuous(
    transform = scales::transform_log10(),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
  ) +
  ggplot2::coord_cartesian(
    ylim = c(10e3, 10e5)
  ) +
  ggplot2::labs(
    title = "Taxa per source type",
    x = "",
    y = "Number of taxa"
  ) +
  ggplot2::guides(
    fill = "none"
  ) +
  ggplot2::scale_fill_manual(
    values = palette_dataset_source_type
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = dataset_source_type,
      xend = dataset_source_type,
      yend = 0
    ),
    color = col_blue_dark,
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      fill = dataset_source_type
    ),
    size = point_size,
    shape = 21,
    col = col_blue_dark,
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_label(
    mapping = ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = point_size * 2,
    position = ggplot2::position_dodge(0.5)
  )

save_figure_to_render(fig_n_taxa_per_source_type)
```

### Classification

To classify the diverse taxa present in the **VegVault** database, the {[taxospace](https://github.com/OndrejMottl/taxospace)} R package was used. This tool automatically aligns taxa names with the [Taxonomy Backbone](https://www.gbif.org/dataset/d7dddbf4-2cf0-4f39-9b2a-bb099caae36c) from the [Global Biodiversity Information Facility](https://www.gbif.org/), providing a standardized classification framework. Specifically, we try to find the best match of the raw names of taxa using [Global Names Resolver](https://resolver.globalnames.org/). Classification information, detailed up to the family level, is stored for each taxon, ensuring consistency and facilitating comparative analyses across different datasets. Currently, the **VegVault** database holds over 110 thousand taxonomic names, of which we were unable to classify only 1312 (1.2%).

```{r}
#| label: TaxonClassification scheme structure
knitr::include_graphics(
  here::here(
    "Technical_overview/DB_scheme_visualisation/TaxonClassification.png"
  )
)
```

```{r}
#| label: taxa classification plot
get_n_taxa_per_level <- function(sel_level) {
  dplyr::tbl(con, "TaxonClassification") %>%
    dplyr::distinct({{ sel_level }}) %>%
    dplyr::collect() %>%
    nrow() %>%
    return()
}

n_taxa_total <-
  dplyr::tbl(con, "Taxa") %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  purrr::chuck("N")

n_taxa_not_classificed <-
  dplyr::left_join(
    dplyr::tbl(con, "Taxa"),
    dplyr::tbl(con, "TaxonClassification"),
    by = "taxon_id"
  ) %>%
  dplyr::filter(
    is.na(taxon_species) &
      is.na(taxon_genus) &
      is.na(taxon_family)
  ) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  purrr::chuck("N")

n_taxa_per_class <-
  tibble::tribble(
    ~class, ~N,
    "total taxa", n_taxa_total,
    "species", get_n_taxa_per_level(taxon_species),
    "genus", get_n_taxa_per_level(taxon_genus),
    "family", get_n_taxa_per_level(taxon_family),
    "not classified", n_taxa_not_classificed
  )

fig_n_taxa_per_class <-
  n_taxa_per_class %>%
  dplyr::mutate(
    class = factor(
      class,
      levels = n_taxa_per_class$class
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = class
    )
  ) +
  ggplot2::labs(
    title = "Taxa classification",
    x = "",
    y = "Number of taxa"
  ) +
  ggplot2::guides(
    fill = "none"
  ) +
  ggplot2::scale_fill_manual(
    values = c(
      "#FCFCFC",
      "#FFE6F5",
      "#FFDEC9",
      "#FFE593",
      "#D8DBE0"
    )
  ) +
  ggplot2::coord_cartesian(
    ylim = c(0, 120e3)
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = class,
      xend = class,
      yend = 0
    ),
    color = col_blue_dark,
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      fill = class
    ),
    size = point_size,
    shape = 21,
    col = col_blue_dark,
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_label(
    mapping = ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = point_size * 2,
    position = ggplot2::position_dodge(0.5)
  )

save_figure_to_render(fig_n_taxa_per_class)
```