---
title: "Section I: Description of VegVaul"
format:
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    self_contained: true
    embed-resources: true
  pdf:
    toc: true
    toc-depth: 3
    fig-width: 10
  gfm:
    toc: true
    toc-depth: 3
    fig-width: 10
    standalone: true
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| label: chunk setup
#| include: FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.path = "figures/",
  out.width = 10
)
```

```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("R/03_Technical_overview/_01_section_I.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Technical_overview/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

**VegVault** is a SQLite interdisciplinary database linking plot-based vegetation data with functional traits and climate. Specifically, it contains:

 - current vegetation plot data
 - past vegetation plot data (fossil pollen records)
 - functional trait data
 - current abiotic data (climate, soil)
 - past abiotic data (climate)

The goal of the database is to compilate interdisciplinary data ...