---
format:
  html:
    fig-width: 10
  gfm:
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("R/03_Technical_overview/_06_section_II_abiotic.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Technical_overview/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

## Abiotic data

Abiotic data is aimed to provide information about all relevant abiotic information affecting vegetation distribution and its traits.

Abiotic data is linked to the structure of the **VegVault** Database by the `gridpoints`, which are artificially created points to *reasonably* cover the resolution of both modern and past data for vegetation and abiotic data.

```{r}
#| label: distribution of gridpoints
x_lim <- c(12, 18.9)
y_lim <- c(48.5, 51.1)

data_grid_coord <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::filter(dataset_type == "gridpoints") %>%
  # this needs to be hardcoded as SQL cannot cope with code imputed values
  dplyr::filter(
    coord_long >= 12 & coord_long <= 18.9
  ) %>%
  dplyr::filter(
    coord_lat >= 48.5 & coord_lat <= 51.1
  )


data_grid_coord %>%
  dplyr::collect() %>%
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = coord_long,
      y = coord_lat
    )
  ) +
  ggplot2::coord_quickmap(
    xlim = x_lim,
    ylim = y_lim
  ) +
  ggplot2::labs(
    title = "Example of gridpoints resolution",
    x = "Longtitude",
    y = "Latitude"
  ) +
  ggplot2::borders(
    fill = NA,
    col = "red"
  ) +
  ggplot2::geom_point(
    shape = 20,
    size = 0.5,
    alpha = 0.75
  )

```


There are currently abiotic from [CHELSA](https://chelsa-climate.org/) and [CHELSA-TRACE21](https://chelsa-climate.org/chelsa-trace21k/) and [WoSIS](https://www.isric.org/explore/wosis). CHELSA and CHELSA-TRACE21 are built on the same structure of variables (visit the websites for more info).

```{r}
#| label: show abitotic variables
dplyr::tbl(con, "AbioticVariable") %>%
  dplyr::select(abiotic_variable_name, abiotic_variable_unit, measure_details) %>%
  dplyr::collect() %>%
  rlang::set_names(
    nm = c(
      "Variable name",
      "Variable unit",
      "source of data"
    )
  ) %>%
  tinytable::tt()
```

Abiotic data is simply linked to `Samples`.

```{r}
#| label: AbioticData scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/AbioticData.png"
  )
)
```

```{r}
#| label: exampe of abitic data - bio1
data_bio1_example <-
  data_grid_coord %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples") %>%
      dplyr::filter(
        age == 0 | (age > 500 & age < 550) |
          (age > 7500 & age < 8500)
      ),
    by = "sample_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "AbioticData") %>%
      dplyr::filter(abiotic_variable_id == 1),
    by = "sample_id"
  ) %>%
  dplyr::collect()

data_bio1_example %>%
  dplyr::select(coord_long, coord_lat, age, abiotic_value) %>%
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = coord_long,
      y = coord_lat
    )
  ) +
  ggplot2::coord_quickmap(
    xlim = x_lim,
    ylim = y_lim
  ) +
  ggplot2::labs(
    title = stringr::str_wrap("Example Mean Annual temprateture at diferent age slices (cal yr BP)", fig_width_def),
    x = "Longtitude",
    y = "Latitude",
    col = "Temperature (°C)",
    fill = "Temperature (°C)"
  ) +
  ggplot2::facet_wrap(~age, nrow = 3) +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::scale_color_viridis_c() +
  ggplot2::theme(
    legend.position = "bottom"
  ) +
  ggplot2::borders(
    fill = NA,
    col = "red"
  ) +
  ggplot2::geom_tile(
    mapping = ggplot2::aes(
      fill = abiotic_value,
      col = abiotic_value
    ),
    alpha = 0.75,
    linewidth = 0.1
  )

```

Note that the spatial resolution is higher for modern climate data than for the past. This is to reduce the size of the past climate data.

```{r}
#| label: exampe of abitic data - soil
data_soil_example <-
  data_grid_coord %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples") %>%
      dplyr::filter(age == 0),
    by = "sample_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "AbioticData") %>%
      dplyr::filter(abiotic_variable_id == 8),
    by = "sample_id"
  ) %>%
  dplyr::collect()

data_soil_example %>%
  dplyr::select(coord_long, coord_lat, age, abiotic_value) %>%
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      x = coord_long,
      y = coord_lat
    )
  ) +
  ggplot2::coord_quickmap(
    xlim = x_lim,
    ylim = y_lim
  ) +
  ggplot2::labs(
    title = "Example of soil types",
    x = "Longtitude",
    y = "Latitude",
    col = "Soit type ID",
    fill = "Soil type ID"
  ) +
  ggplot2::facet_wrap(~age, nrow = 3) +
  ggplot2::theme(
    legend.position = "none"
  ) +
  ggplot2::borders(
    fill = NA,
    col = "red"
  ) +
  ggplot2::geom_tile(
    mapping = ggplot2::aes(
      fill = as.factor(abiotic_value),
      col = as.factor(abiotic_value)
    ),
    alpha = 0.5,
    linewidth = 0.2
  )
```