---
format:
  html:
    fig-width: 10
  gfm:
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("R/03_Technical_overview/_03_section_II_samples.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Technical_overview/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

## Samples

`Sample` represents the main unit of data in the **VegVault** database. 

```{r}
#| label: Samples scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/Samples.png"
  )
)
```

### Dataset-Sample

First `Samples` are linked to `Datasets` via the `Dataset-Sample` table.

```{r}
#| label: DatasetSample scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/DatasetSample.png"
  )
)
```

```{r}
#| label: Number of samples plot
n_samples_per_dataset_type <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples"),
    by = "sample_id"
  ) %>%
  dplyr::distinct(dataset_type, sample_id) %>%
  dplyr::group_by(dataset_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

plot_waffle(
  n_samples_per_dataset_type,
  dataset_type,
  plot_title = "Number of samples per Dataset type",
  one_point_is = 25e3
) +
  ggplot2::scale_fill_manual(
    values = palette_dataset_type
  )

plot_waffle(
  n_samples_per_dataset_type %>%
    dplyr::filter(
      dataset_type != "gridpoints"
    ),
  dataset_type,
  plot_title = "Number of samples per Dataset type - no gridpoints",
  one_point_is = 25e3
) +
  ggplot2::scale_fill_manual(
    values = palette_dataset_type
  )
```

### Sample-size

Vegetation plots can have different sizes, which can have a huge impact on analyses. Therefore, the information about the plot is saved separately. 

```{r}
#| label: SampleSizeID scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/SampleSizeID.png"
  )
)
```

```{r}
#| label: plot-size plot
plot_size_types <-
  dplyr::tbl(con, "SampleSizeID") %>%
  dplyr::distinct(description) %>%
  dplyr::collect()

if (
  nrow(plot_size_types) == 1
) {
  data_samples_plot_size <-
    dplyr::inner_join(
      dplyr::tbl(con, "Datasets"),
      dplyr::tbl(con, "DatasetTypeID"),
      by = "dataset_type_id"
    ) %>%
    dplyr::filter(
      dataset_type == "vegetation_plot"
    ) %>%
    dplyr::inner_join(
      dplyr::tbl(con, "DatasetSample"),
      by = "dataset_id"
    ) %>%
    dplyr::inner_join(
      dplyr::tbl(con, "Samples"),
      by = "sample_id"
    ) %>%
    dplyr::left_join(
      dplyr::tbl(con, "SampleSizeID"),
      by = "sample_size_id"
    ) %>%
    dplyr::select(sample_size) %>%
    dplyr::collect()

  data_samples_plot_size %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = sample_size,
      )
    ) +
    ggplot2::scale_x_continuous(
      trans = "log10"
    ) +
    ggplot2::scale_y_continuous(
      trans = "log10",
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    ggplot2::labs(
      title = "Vegetation plot size",
      x = paste0("plot size (", plot_size_types$description, ")"),
      y = "Number of Samples"
    ) +
    ggplot2::geom_histogram(
      fill = "grey50",
      col = NA,
      bins = 15
    )
}
```

### Sample age

The **VegVault** database deals with both current and paleo data. Therefore, each `Sample` has the indication of *age*, with modern samples being set to 0. To embrace the uncertainty from age-depth modeling paleo-record, the **VegVault** database has a structure to hold an uncertainty matrix containing information about all *potential ages* of each `Sample` from a paleo `Dataset`. 

```{r}
#| label: SampleUncertainty scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/SampleUncertainty.png"
  )
)
```

We can show this on the previously selected fossil pollen archive with dataset ID: `r as.character(sel_dataset_for_example)`.

```{r}
#| label: Sample poential age - plot
dataset_age_example <-
  dplyr::tbl(con, "Datasets") %>%
  dplyr::filter(dataset_id == sel_dataset_for_example) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples"),
    by = "sample_id"
  ) %>%
  dplyr::distinct(sample_id, age) %>%
  dplyr::collect()

dataset_age_example_iterations <-
  dplyr::tbl(con, "Datasets") %>%
  dplyr::filter(dataset_id == sel_dataset_for_example) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSample"),
    by = "dataset_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "Samples"),
    by = "sample_id"
  ) %>%
  dplyr::distinct(sample_id) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "SampleUncertainty"),
    by = "sample_id"
  ) %>%
  dplyr::select(
    sample_id,
    age_it = age
  ) %>%
  dplyr::collect()

dplyr::left_join(
  dataset_age_example_iterations,
  dataset_age_example,
  by = "sample_id"
) %>%
  ggplot2::ggplot(
    mapping = ggplot2::aes(
      y = age,
      group = as.factor(age),
      x = age_it
    )
  ) +
  ggplot2::coord_fixed() +
  ggplot2::labs(
    title = paste0(
      "Dataset ID:",
      as.character(sel_dataset_for_example)
    ),
    y = "Sample age (cal yr BP)",
    x = "Potential age of Samples (cal yr BP)"
  ) +
  ggplot2::geom_boxplot(
    outlier.shape = NA
  )

```

### Sample reference

Individual `Sample` can have specific references on top of the reference to `Dataset`

```{r}
#| label: SampleReference scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/SampleReference.png"
  )
)
```