---
format:
  html:
    fig-width: 10
  gfm:
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---
```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("R/03_Technical_overview/_05_section_II_traits.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Technical_overview/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

## Traits

Functional traits of vegetation taxa follow the same structure of `Dataset` and `Samples` obtained directly from *Dataset Source-types*.

```{r}
#| label: Traits scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/Traits.png"
  )
)
```

### Trait domain

As there are many varying names for the same "traits", the **VegVault** database contains *Trait Domain* information to group traits together.  

```{r}
#| label: TraitsDomain scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/TraitsDomain.png"
  )
)
```

There are currently 6 trait domains following the [Diaz et al. (2016)](https://www.nature.com/articles/nature16489)

```{r}
#| label: show trait domains
dplyr::tbl(con, "TraitsDomain") %>%
  dplyr::distinct(trait_domain_name) %>%
  dplyr::collect() %>%
  rlang::set_names(
    nm = "Trait domain"
  ) %>%
  tinytable::tt()
```

```{r}
#| label: trait per domain plot

n_traits_per_domain <-
  dplyr::inner_join(
    dplyr::tbl(con, "TraitsDomain"),
    dplyr::tbl(con, "Traits"),
    by = "trait_domain_id"
  ) %>%
  dplyr::distinct(trait_domain_name, trait_name) %>%
  dplyr::group_by(trait_domain_name) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

n_traits_per_domain %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = trait_domain_name
    )
  ) +
  ggplot2::labs(
    title = "Trait domains",
    x = "",
    y = "Number of traits"
  ) +
  ggplot2::guides(
    fill = "none"
  ) +
  ggplot2::scale_fill_manual(
    values = palette_trait_dommanins
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = trait_domain_name,
      xend = trait_domain_name,
      yend = 0
    ),
    color = "grey",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      fill = trait_domain_name
    ),
    size = 3,
    shape = 21,
    col = "black",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_label(
    mapping = ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = 3,
    position = ggplot2::position_dodge(0.5)
  )

```

### Trait Values

To store a trait value, information needs to be linked among `Dataset`, `Sample`, `Taxa`, and `Trait`.

```{r}
#| label: TraitsValue scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/TraitsValue.png"
  )
)
```

```{r}
#| label: trait value occurences plot

n_occurences_per_dommain <-
  dplyr::inner_join(
    dplyr::tbl(con, "TraitsDomain"),
    dplyr::tbl(con, "Traits"),
    by = "trait_domain_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "TraitsValue"),
    by = "trait_id"
  ) %>%
  dplyr::group_by(trait_domain_name) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

n_occurences_per_dommain %>%
  dplyr::arrange(-N) %>%
  dplyr::mutate(
    trait_domain_name = factor(
      trait_domain_name,
      levels = unique(trait_domain_name)
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = trait_domain_name
    )
  ) +
  ggplot2::scale_y_continuous(
    transform = scales::transform_log10(),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  ggplot2::labs(
    title = "Occurences of trait values",
    x = "",
    y = "Number of occurrences (values)"
  ) +
  ggplot2::guides(
    fill = "none"
  ) +
  ggplot2::scale_fill_manual(
    values = palette_trait_dommanins
  ) +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = trait_domain_name,
      xend = trait_domain_name,
      yend = 0
    ),
    color = "grey",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      fill = trait_domain_name
    ),
    size = 3,
    shape = 21,
    col = "black",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_label(
    mapping = ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = 3,
    position = ggplot2::position_dodge(0.5)
  )
```

### Trait reference

For full clarity, on top of `Dataset` and `Sample`, each `Trait` can have additional references.

```{r}
#| label: TraitsReference scheme structure
knitr::include_graphics(
  here::here(
    "R/03_Technical_overview/DB_scheme_visualisation/TraitsReference.png"
  )
)
```