---
title: "EDA of the VegVault database"
date: last-modified
author: Ondřej Mottl
output:
  html_document:
    toc: yes
    toc_float: yes
    fig-width: 10
    self_contained: yes
  pdf_document:
    toc: yes
    toc_depth: 3
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---

# VegVault database

## Section I: Description of VegVaul

VegVault is a SQLite interdisciplinary database linking plot-based vegation data with functinal traits and climate. Specifically, it contains:

 - current vegetation plot data
 - past vegetation plot data (fossil pollen reocrds)
 - trait data
 - current abiotic data (climate, soil)
 - past abiotic data (climate)

The goal of the 


## Section II: Overview of VegVault 1.0.0

```{r}
#| label: chunk setup
#| include: FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.path = "figures/"
)
```


```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("R/03_Supplementary_analyses/01_DB_EDA.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

```{r}
#| label: chunk theme
#| include: FALSE
ggplot2::theme_set(
  ggplot2::theme_bw() +
    ggplot2::theme(
      axis.title = ggplot2::element_text(size = 25),
      axis.text = ggplot2::element_text(size = 15),
      strip.text = ggplot2::element_text(size = 15),
      panel.grid = ggplot2::element_blank()
    )
)

fig_width_def <- 60 # this is used to wrap text.

#' @description
#' A helper function to colour the facets
color_facets <-
  function(sel_plot,
           sel_palette,
           direction = c("vertical", "horizontal"),
           return_raw = FALSE) {
    direction <- match.arg(direction)
    g <-
      ggplot2::ggplot_gtable(
        ggplot2::ggplot_build(sel_plot)
      )
    stripr <-
      which(grepl("strip-t", g$layout$name))

    for (i in seq_along(stripr)) {
      obejct_val <-
        sort(stripr,
          decreasing = ifelse(direction == "vertical",
            TRUE,
            FALSE
          )
        )[i]

      j <-
        which(grepl("rect", g$grobs[[obejct_val]]$grobs[[1]]$childrenOrder))

      g$grobs[[obejct_val]]$grobs[[1]]$children[[j]]$gp$fill <-
        sel_palette[i]
    }

    if (
      return_raw == TRUE
    ) {
      return(g)
    } else {
      grid::grid.draw(g)
    }
  }
```

```{r}
# |label: connect to DB

con <-
  DBI::dbConnect(
    RSQLite::SQLite(),
    paste0(
      data_storage_path,
      "Data/VegVault/VegVault.sqlite"
    )
  )
```

```{r}
#----------------------------------------------------------#
# 2. Datasets -----
#----------------------------------------------------------#

n_datasetes <-
  dplyr::tbl(con, "Datasets") %>%
  dplyr::count() %>%
  dplyr::collect() %>%
  purrr::chuck("n")

n_datasetes_per_type <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::group_by(dataset_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

plot_waffle(
  n_datasetes_per_type,
  dataset_type,
  plot_title = "Datasets type",
  one_point_is = 1e3,
  n_rows = 50
)

n_datasetes_per_type %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = reorder(dataset_type, -N)
    )
  ) +
  ggplot2::scale_y_continuous(
    trans = "log10",
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  ggplot2::labs(
    title = "Datasets type",
    x = "",
    y = "Number of datasets"
  ) +
  ggplot2::geom_segment(
    ggplot2::aes(
      xend = dataset_type,
      yend = 0
    ),
    color = "grey"
  ) +
  ggplot2::geom_point(
    size = 3,
    shape = 21,
    col = "black",
    fill = "white"
  ) +
  ggplot2::geom_label(
    ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = 3
  )

n_datasetes_per_source_type <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetSourceTypeID"),
    by = "data_source_type_id"
  ) %>%
  dplyr::group_by(dataset_source_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

plot_waffle(
  n_datasetes_per_source_type,
  dataset_source_type,
  plot_title = "Datasets source type",
  one_point_is = 1e3,
  n_rows = 50
)
```


```{r}
n_datasetes_per_type_per_source_type <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSourceTypeID"),
    by = "data_source_type_id"
  ) %>%
  dplyr::group_by(dataset_type, dataset_source_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()

n_datasetes_per_type_per_source_type %>%
  dplyr::group_by(dataset_type) %>%
  dplyr::mutate(
    n_datasetes_per_type = sum(N)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(-n_datasetes_per_type, -N) %>%
  dplyr::mutate(
    dataset_type = factor(
      dataset_type,
      levels = unique(dataset_type)
    ),
    dataset_source_type = factor(
      dataset_source_type,
      levels = unique(dataset_source_type)
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(
      y = N,
      x = dataset_source_type # reorder(dataset_source_type, -N)
    )
  ) +
  ggplot2::scale_y_continuous(
    trans = "log10",
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  ggplot2::labs(
    title = "Datasets type per source type",
    x = "",
    y = "Number of datasets"
  ) +
  ggplot2::guides(
    fill = "none"
  ) +
  ggplot2::facet_wrap(~dataset_type, nrow = 1, scales = "free_x") +
  ggplot2::geom_segment(
    mapping = ggplot2::aes(
      x = dataset_source_type, # reorder(dataset_source_type, -N),
      xend = dataset_source_type, # reorder(dataset_source_type, -N),
      yend = 0
    ),
    color = "grey",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_point(
    mapping = ggplot2::aes(
      fill = dataset_source_type
    ),
    size = 3,
    shape = 21,
    col = "black",
    position = ggplot2::position_dodge(0.5)
  ) +
  ggplot2::geom_label(
    mapping = ggplot2::aes(
      label = scales::number(N)
    ),
    vjust = -0.5,
    size = 3,
    position = ggplot2::position_dodge(0.5)
  )

n_dataset_sources <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetTypeID"),
    by = "dataset_type_id"
  ) %>%
  dplyr::inner_join(
    dplyr::tbl(con, "DatasetSourcesID"),
    by = "data_source_id"
  ) %>%
  dplyr::distinct(dataset_type, data_source_desc) %>%
  dplyr::group_by(dataset_type) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()


View(n_dataset_sources)

n_datasetes_per_source <-
  dplyr::inner_join(
    dplyr::tbl(con, "Datasets"),
    dplyr::tbl(con, "DatasetSourcesID"),
    by = "data_source_id"
  ) %>%
  dplyr::group_by(data_source_desc) %>%
  dplyr::count(name = "N") %>%
  dplyr::collect() %>%
  dplyr::ungroup()


plot_waffle(
  n_datasetes_per_source,
  data_source_desc,
  plot_title = "Datasets source",
  one_point_is = 1e3,
  n_rows = 50
)
```

## Section III: Assebly details of VegVault 1.0.0