---
format:
  html:
    fig-width: 10
editor: 
  markdown: 
    wrap: sentence
---

# Database content

The **VegVault** database v1.0.0 is an SQLite database of a size of ~ 110 GB

Currently, the `SQLite` database consists of 29 interconnected tables with 87 fields (variables). See [Databse Structure](/website/database_structure.qmd) for more information.

|||
|-----|-----|
Datasets | > 480 000 |
Samples | > 13 000 000|
Taxa | > 110 000|
Vegetation Traits | 6|
Trait values | > 11 000 000|
Abiotic variables | 8|


See [Database Access](/website/database_access.qmd) for more information on how to access the database.

## Assembly details of VegVault 1.0.0

```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("website/database_content.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Visualisation/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

The **VegVault** version **1.0.0** has been assembled from other publicly available databases. 


## Assembly process

Here is a general visualisation of the assembly process:

```{r}
#| label: VV assembly scheme
knitr::include_graphics(
  here::here(
    "website/database_structure/Figures/DB_assembly_visualisation/assembly_viz_01.png"
  )
)
```

**Figure legend**:

1. **[Neotoma Paleoecology Database](https://www.neotomadb.org/)** - open, community-curated data and services for paleoecological and paleoenvironmental data
     * source of fossil pollen data 
2. **[sPlotOpen](https://idiv-biodiversity.de/en/splot/splotopen.html)** - The open-access version of sPlot
     * source of current vegetation plot data
3. **[Botanical Information and Ecology Network](https://bien.nceas.ucsb.edu/bien/)** - global patterns of plot inventories and surveys
    * source of current vegetation plot and vegetation trait data
4.  **[TRY Plant Trait Database](https://www.try-db.org/TryWeb/Home.php)** - open access plant trait data 
    * source of vegetation trait data
5. **[Climatologies at High resolution for the Earth's Land Surface Area](https://chelsa-climate.org/)** - Long-term, transient modern- and paleo-climate data
    * source of abiotic data
6. **[World Soil Information Service](https://www.isric.org/explore/wosis)** - harmonised soil profile database
    * source of abiotic data
7.  **[FOSSILPOL](https://hope-uib-bio.github.io/FOSSILPOL-website/)** - The workflow that aims to process and standardise global palaeoecological pollen data for macroecological studies
    * a method to obtain and process fossil pollen data
8.  **[RBIEN R package](https://github.com/bmaitner/RBIEN)** - Tools for accessing the BIEN database
    * a method to obtain current vegetation plot data 
9.  **[VegVault-FOSSILPOL GitHub repo](https://github.com/OndrejMottl/VegVault-FOSSILPOL/tree/v1.0.0)**
    * a Tag (v1.0.0) to obtain and process fossil pollen data
10. **[VegVault-Vegetation_data GitHub repo](https://github.com/OndrejMottl/VegVault-Vegetation_data/tree/v1.0.0)**
    * a Tag (v1.0.0) to process current vegetation plot data
11. **[VegVault-Trait_data GitHub repo](https://github.com/OndrejMottl/VegVault-Trait_data/tree/v1.1.0)**
    * a Tag (v1.1.0) to process vegetation trait data
12. **[VegVault-abiotic_data GitHub repo](https://github.com/OndrejMottl/VegVault-abiotic_data/tree/v1.1.0)**
    * a Tag (v1.1.0) to process abiotic data
13. **[VegVault GitHub repo](https://github.com/OndrejMottl/VegVault)**
    * a Tag (v1.0.0) to transfer the data into the SQLite database
14. **VegVault** - SQLite database (v1.0.0)

### Note on GitHub Tags

*...Tags are ref's that point to specific points in Git history. Tagging is generally used to capture a point in history...* 

Here, the Tags have been used to ensure reproducibility of this version of **VegVault**. GitHub repo can change but the specific Tag used here (and also used in the code to source the data) ensure the transparency of the state of the database. 

