---
date: 2025/01/24
date-format: long
date-modified: last-modified
format:
  html:
    fig-width: 10
sidebar: true
---

# Database content

The **VegVault** database v1.0.0 is an SQLite database of a size of ~ 110 GB

Currently, the `SQLite` database consists of 29 interconnected tables with 87 fields (variables). See [Databse Structure](/website/database_structure.qmd) for more information.

|||
|-----|-----|
Datasets | > 480 000 |
Samples | > 13 000 000|
Taxa | > 110 000|
Vegetation Traits | 6|
Trait values | > 11 000 000|
Abiotic variables | 8|


See [Database Access](/website/database_access.qmd) for more information on how to access the database.


!!! There should be spatio-temporal overview of the database content !!!

## Assembly details of VegVault 1.0.0

```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("website/database_content.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Visualisation/00_Config_quarto.R"
  )
)

```

The **VegVault** version **1.0.0** has been assembled from other publicly available databases. 


## Assembly process

Here is a general visualisation of the assembly process:

```{r}
#| label: VV assembly scheme
knitr::include_graphics(
  here::here(
    "Outputs/Figures/DB_assembly_visualisation/assembly_viz_04.png"
  )
)

```

**Figure legend**:

1. **[Neotoma Paleoecology Database](https://www.neotomadb.org/)** - open, community-curated data and services for paleoecological and paleoenvironmental data
     * source of fossil pollen data 
2. **[sPlotOpen](https://idiv-biodiversity.de/en/splot/splotopen.html)** - The open-access version of sPlot
     * source of current vegetation plot data
3. **[Botanical Information and Ecology Network](https://bien.nceas.ucsb.edu/bien/)** - global patterns of plot inventories and surveys
    * source of current vegetation plot and vegetation trait data
4.  **[TRY Plant Trait Database](https://www.try-db.org/TryWeb/Home.php)** - open access plant trait data 
    * source of vegetation trait data
5. **[Climatologies at High resolution for the Earth's Land Surface Area](https://chelsa-climate.org/)** - Long-term, transient modern- and paleo-climate data
    * source of abiotic data
6. **[World Soil Information Service](https://www.isric.org/explore/wosis)** - harmonised soil profile database
    * source of abiotic data
7.  **[VegVault-FOSSILPOL GitHub repo](https://github.com/OndrejMottl/VegVault-FOSSILPOL/tree/v1.0.0)**
    * a Tag (v1.0.0) to obtain and process fossil pollen data
8.  **[VegVault-Vegetation_data GitHub repo](https://github.com/OndrejMottl/VegVault-Vegetation_data/tree/v1.0.0)**
    * a Tag (v1.0.0) to process current vegetation plot data
9.  **[VegVault-Trait_data GitHub repo](https://github.com/OndrejMottl/VegVault-Trait_data/tree/v1.1.0)**
    * a Tag (v1.1.0) to process vegetation trait data
10. **[VegVault-abiotic_data GitHub repo](https://github.com/OndrejMottl/VegVault-abiotic_data/tree/v1.1.0)**
    * a Tag (v1.1.0) to process abiotic data
11. **[VegVault GitHub repo](https://github.com/OndrejMottl/VegVault)**
    * a Tag (v1.0.0) to transfer the data into the SQLite database. S
    * ee details below.
12. **VegVault** - SQLite database (v1.0.0)

### Note on GitHub Tags

*...Tags are ref's that point to specific points in Git history. Tagging is generally used to capture a point in history...* 

Here, the Tags have been used to ensure reproducibility of this version of **VegVault**. GitHub repo can change but the specific Tag used here (and also used in the code to source the data) ensure the transparency of the state of the database. 

### VegVault GitHub repo

In addition to the consolidation of all processed data into a unified SQLite database, the final **VegVault** "migration" repository performs three additional procedures: (i) taxa classification, (ii) grouping of traits into Trait Domains, and (iii) creation of gridpoints for abiotic data. 

* #### (i) Taxa classification 
  As **VegVault** consist of data on taxa from various sources, the [{taxospace} R package](https://github.com/OndrejMottl/taxospace) is used to classify the diverse taxa into a unifying backbone in the **VegVault** database. The {taxospace} tool automatically aligns taxa names with the taxonomical backbone of the GBIF. Specifically, we find the best match of the raw names of taxa using [Global Names Resolver](https://resolver.globalnames.org/)  that is then aligned with GBIF. Taxonomic information, detailed up to the family level, is stored for each taxon, ensuring consistency and facilitating comparative analyses across different datasets.
* #### (ii) Grouping of Traits into Trait Domains 
  As there are differences in trait names across sources of data (e.g., "*Leaf nitrogen (N) content per leaf dry mass*" and "*leaf nitrogen content per leaf dry mass*"), we added a new variable `Trait Domain` that   groups traits together following the trait selection of Diaz et al. (2016). For example, trait "*Plant height vegetative*" from TRY and trait "*whole plant height*" from BIEN are both grouped under "*Plant height*" `Trait Domain`. 
* #### (iii) Creation of gridpoints of abiotic data
  The abiotic data in the **VegVault** database provide essential information on environmental factors affecting vegetation distribution and traits, namely climate and soil variables. We developed a data structure that provides readily available environmental context for each vegetation (and trait) record by creating spatio-temporal links between these records and abiotic information. We first unified the resolution of the original rasters by resampling all abiotic data into ~ 25km resolution at 500-year time slices, reducing the amount of data. Next, as the rasters are not suitable to be stored in an SQLite database, we created artificial points, called `gridpoints`, located in the middle of each raster cell. This resulted in a uniform spatio-temporal matrix of `gridpoints` holding the abiotic information. 
  Next, we linked `gridpoints` and other non-gridpoint `Samples`, namely vegetation_plot, fossil_pollen_archive, and traits (see [database structure](/website/database_structure.qmd)), and calculated the spatial and temporal distances between them. We discarded any gridpoint `Sample` beyond 50 km and/or 5000 years to any other non-gridpoint `Sample`. In **VegVault**, users can select the information for each non-gridpoint `Sample` from the closest spatio-temporal abiotic data or get the average from all surrounding `gridpoints`. By providing such comprehensive and well-linked structure between vegetation and abiotic data, **VegVault** enhances the ability to study the interactions between vegetation and their environment, facilitating the workflow to advanced ecological research and modelling efforts. 
