---
title: "Examples of VegVault usage"
format:
  html:
    fig-width: 10
date: 2025/01/24
date-format: long
date-modified: last-modified
sidebar: true
---

```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("website/_database_examples.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)

source(
  here::here(
    "R/03_Visualisation/00_Config_quarto.R"
  )
)

invisible(
  lapply(
    list.files(
      path = here::here("R/Functions"),
      pattern = "*.R",
      recursive = TRUE,
      full.names = TRUE
    ),
    source
  )
)
```

:::: {.columns}
::: {.column width="75%"}
In order to make usage of the **VegVault** database as easy as possible, we have developed an R-package called [{vaultkeepr}](https://bit.ly/vaultkeepr), providing a suite of functions to effectively interact with the database directly from the [R programming language](https://en.wikipedia.org/wiki/R_(programming_language)). Functions include opening the database, extracting datasets, filtering samples, and accessing specific taxa and traits. This package is a well-tested interface (>95% code coverage).
:::

::: {.column width="25%"}
```{r}
#| label: vaultkeepr-logo
knitr::include_graphics(
  "https://ondrejmottl.github.io/vaultkeepr/reference/figures/vaultkeeper_logo.png"
)
```
:::

::::

The {vaultkeepr} can be installed from [GitHub](https://github.com/) with:

```{r}
#| label: install
#| eval: false
#| echo: true
# install.packages("remotes")
remotes::install_github("OndrejMottl/vaultkeepr")
```

and then all functions will be made available by attaching as: 

```{r}
#| label: example
#| echo: true
library(vaultkeepr)
```

Here we present 3 examples of potential projects and how to obtain data for such projects from **VegVault** using {vaultkeepr} R-package. Note that we are specifically not trying to do any analysis, only presenting a way to obtain data, which can be used for such projects.

{{< include _database_examples.md >}}