---
date: 2025/01/24
date-format: long
date-modified: last-modified
sidebar: true
---

# Database Access & Usage

## Download VegVault Database

[VegVault]{.text-background-black .text-color-white .text-bold} v1.0.0 is available as a comprehensive [SQLite database]{.text-background-brownDark .text-color-white .text-bold} providing integrated access to global vegetation, [trait]{.text-background-blueDark .text-color-white .text-bold}, and [environmental data]{.text-background-greenDark .text-color-white .text-bold}.

[Download VegVault v1.0.0 (SQLite)](https://doi.org/10.48700/datst.t21cz-2nq071){.btn} 

::: {.callout-note}
**System Requirements**: [VegVault]{.text-background-black .text-color-white .text-bold} contains over 100GB of data. Ensure your system has sufficient storage space and memory (RAM) before downloading.
:::

### Alternative Access: Flat Files

For users who prefer working with individual data tables, we provide [VegVault]{.text-background-black .text-color-white .text-bold} as separate CSV files:

[Download VegVault v1.0.0 (Flat Files)](https://doi.org/10.48700/datst.thktf-5rr35){.btn} 

::: {.callout-warning}
**Note**: Flat files are not the [recommended]{.text-background-brownLight .text-color-white .text-bold} approach. Many files exceed 1GB and may be difficult to load into memory. We strongly recommend using the [SQLite database]{.text-background-brownDark .text-color-white .text-bold} with the {vaultkeepr} R package for optimal performance.
:::

## Accessing VegVault with {vaultkeepr}

:::: {.columns}
::: {.column width="75%"}
To make [VegVault]{.text-background-black .text-color-white .text-bold} easily accessible, we developed [{vaultkeepr}](https://bit.ly/vaultkeepr), a well-tested R package (>95% code coverage) that provides intuitive functions for interacting with the [database]{.text-background-brownDark .text-color-white .text-bold}. The package handles database connections, data filtering, and extraction, allowing you to retrieve only project-specific data without loading everything into memory.
:::

::: {.column width="25%"}
```{r}
#| label: vaultkeepr-logo
#| echo: false
knitr::include_graphics(
  "https://ondrejmottl.github.io/vaultkeepr/reference/figures/vaultkeeper_logo.png"
)
```
:::
::::

### Installation

Install {vaultkeepr} from GitHub:

```{r}
#| label: install
#| eval: false
#| echo: true
# install.packages("remotes")
remotes::install_github("OndrejMottl/vaultkeepr")
```

Load the package:

```{r}
#| label: example
#| echo: true
library(vaultkeepr)
```

### Quick Start

See our [Usage Examples](/website/database_examples.qmd) for detailed demonstrations of how to:

- Extract vegetation data for specific regions and time periods
- Access functional trait information
- Link abiotic environmental data
- Perform taxonomic harmonization

## Guidelines for Data Reuse and Citation

### Licensing

VegVault is released under **CC BY 4.0 License** ([Creative Commons Attribution 4.0](https://creativecommons.org/licenses/by/4.0/)), allowing you to:

- Copy and redistribute the material
- Adapt and build upon the material

**Requirements**: You must provide appropriate attribution to VegVault and cite all primary data sources.

### Citation Requirements

When using VegVault in your research:

1. **Cite VegVault**: Always cite the latest VegVault version with its corresponding DOI
2. **Cite primary sources**: Acknowledge all original data providers included in your analysis
3. **Use our helper function**: The `vaultkeepr::get_references()` function automatically generates the complete list of recommended citations for your specific data compilation

### Publication & Pre-print

**Pre-print**: "*VegVault database: linking global paleo-, and neo-vegetation data with functional traits and abiotic drivers*" 

[Access Pre-print](https://doi.org/10.1101/2025.01.31.635213){.btn}

For additional citation materials, visit our [Materials & Outputs](/website/materials.qmd) page.

## Important Limitations

### Data Quality Considerations

VegVault's quality depends on the completeness, accuracy, and reliability of primary data sources. We acknowledge potential issues in global ecological databases (as noted for [TRY](https://www.try-db.org/)) but focus on improving data re-usability rather than correcting source data quality issues.

### Technical Limitations

- **Storage requirements**: >100GB database size may be challenging for some systems
- **Memory requirements**: Large datasets may require substantial RAM for analysis
- **Version dependencies**: Always specify the VegVault version used in your research

## Additional Resources

- **[Database Structure](/website/database_structure.qmd)** - Detailed technical documentation
- **[Database Content](/website/database_content.qmd)** - Overview of included data
- **[Database Assembly](/website/database_assembly.qmd)** - Data processing and integration procedures
- **[Database Examples](/website/database_examples.qmd)** - Practical usage scenarios
- **[Contact & Support](/website/contact.qmd)** - Get help and report issues