---
date: 2025/01/24
date-format: long
date-modified: last-modified
sidebar: true
---

# Materials & Outputs


```{r}
#| label: source config
#| include: FALSE
#| results: 'hide'
#| warning: FALSE
#| message: FALSE

library(here)

here::i_am("website/materials.qmd")

# Load configuration
source(
  here::here(
    "R/00_Config_file.R"
  )
)
```

This page provides access to all VegVault-related resources including the database, publications, code, and presentation materials.

## VegVault Database

:::: {.columns}

::: {.column width="70%"}

### Primary Database Access

**VegVault v1.0.0 (SQLite Database - Recommended)**
[Download Database v1.0.0](https://doi.org/10.48700/datst.t21cz-2nq07){.btn}

Please review our [Database Access Guidelines](database_access.qmd#guidelines-for-data-reuse-and-citation) for proper usage and citation requirements.

:::

::: {.column width="30%"}

```{r}
#| label: vegvault-logo
#| echo: false
knitr::include_graphics(
  here::here(
    "Outputs/Figures/website/VegVault_logo.png"
  )
)
```

:::

::::

### Alternative Format

**VegVault v1.0.0 (Flat Files - Not Recommended)**
[Download Database v1.0.0 as Flat Files](https://doi.org/10.48700/datst.thktf-5rr35){.btn}

::: {.callout-warning}
**Note**: Flat files are provided for accessibility but are not the preferred method. Many files exceed 1GB and may be difficult to load. We strongly recommend using the SQLite database with [{vaultkeepr}](https://bit.ly/vaultkeepr) for optimal performance.
:::

## Publications

### Main Article

**"VegVault dataset: linking global paleo-, and neo-vegetation data with functional traits and abiotic drivers"** [![](https://img.shields.io/badge/DOI-10.1038/s41597--025--06176--1-blue)](https://doi.org/10.1038/s41597-025-06176-1)

### Pre-print

**"VegVault database: linking global paleo-, and neo-vegetation data with functional traits and abiotic drivers"** [![](https://img.shields.io/badge/DOI-10.1101/2025.01.31.635213-blue)](https://doi.org/10.1101/2025.01.31.635213)

## Software & Tools

### {vaultkeepr} R Package

:::: {.columns}

::: {.column width="75%"}

**Official R package for VegVault database interaction**
[{vaultkeepr} Documentation & Download](https://bit.ly/vaultkeepr){.btn}

The package provides:

- Easy database connection and querying
- Data filtering and extraction functions
- Automatic citation generation
- [![Codecov-test-coverage](https://codecov.io/gh/OndrejMottl/vaultkeepr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/OndrejMottl/vaultkeepr?branch=main) code coverage with comprehensive testing
:::

::: {.column width="25%"}
```{r}
#| label: vaultkeepr-logo
#| echo: false
knitr::include_graphics(
  "https://ondrejmottl.github.io/vaultkeepr/reference/figures/vaultkeeper_logo.png"
)
```
:::

::::

## Visual Resources

## Presentations

### Recent Presentations

**VegVault - JCU, České Budějovice, Czech Republic (December 110, 2025)**

- [View Slides](https://ondrejmottl.github.io/VegVault-Presentation-20251210/)
- [Source Code](https://github.com/OndrejMottl/VegVault-Presentation-20251210)

**VegVault - IAVS Conference, Colorado, USA (August 2, 2025)**

- [View Slides](https://ondrejmottl.github.io/VegVault-Presentation-20250802/)
- [Source Code](https://github.com/OndrejMottl/VegVault-Presentation-20250802)

**VegVault Flash Presentation - Geobotany Meeting, Sudějov (March 23, 2025)**

- [View Slides](https://ondrejmottl.github.io/VegVault-FlashPresentation-20250323/)
- [Source Code](https://github.com/OndrejMottl/VegVault-FlashPresentation-20250323)

### Conference Posters

**VegVault Poster - CSPE Conference (September 2024)**

- [Low Resolution Version](/Outputs/Figures/poster2024/mottl_poster_compression_9.png)  
- [High Resolution Version](/Outputs/Figures/poster2024/mottl_poster.jpeg)

### VegVault Logo

[Download VegVault Logo](https://github.com/OndrejMottl/VegVault/tree/main/Outputs/Figures/website/VegVault_logo.png){.btn}

Available for use in presentations, publications, and educational materials with proper attribution.


## Additional Resources

- **[Database Access](database_access.qmd)** - Download and setup instructions
- **[Database Assembly](database_assembly.qmd)** - Data processing and integration procedures
- **[Database Examples](database_examples.qmd)** - Practical usage examples
- **[Database Structure](database_structure.qmd)** - Technical documentation
- **[Contact & Support](contact.qmd)** - Get help and report issues

